---
title: "analyzeData"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lme4)
library(optimx)
library(DHARMa)
library(fuzzyjoin)

mod_human<-read.csv(file = '../Data/temporalMod.csv') %>% 
  select(-1)
mod_phys<-read.csv(file = '../Data/spatialMod.csv') %>% 
  select(-1)
```



## Models
fitted following Zuur (2009) methods for validation
### Spatial Models

####Breeding Pairs
```{r poisson regression}
hist(mod_phys$max_BP)
mean(mod_phys$max_BP, na.rm = TRUE) #1.559
var(mod_phys$max_BP, na.rm = TRUE) #3.044

m1<-glm(max_BP ~ BeachArea_km + beach_width + Length_km + parkArea + bufferPop_km2 + Border_road + tot_area_km_Beach + developedLand + Management + Region, data = mod_phys, family = poisson(link = "log")) # no error

m2<-glm(max_BP ~ BeachArea_km + beach_width + Length_km + parkArea + bufferPop_km2 + Border_road + tot_area_km_Beach + developedLand  + Region, data = mod_phys, family = poisson(link = "log")) # no error

summary(m1)
summary(m2) #better AIC and I'm not happy with management and it's designations right now
```

all good
```{r model validation}
# with DHARMa

simOutput <- simulateResiduals(fittedModel = m2, plot = T)
#some QQ plot deviation but n.s. I'm okay with this model (DHARMa documentation)

#tested zero inflation because we have a lot of 0 presence observations
testZeroInflation(simOutput) #this is okay, n.s. and well within the distribution

#overdispersion is a common issue with poisson
testDispersion(simOutput, type = "PearsonChisq") #okay

```


```{r spatial model effect size}
sjPlot::plot_model(m2, show.values = TRUE, axis.lim = c(.001,1000), sort.est = TRUE, value.offset = .3)

```


#### Productivity models

note productivity model is zero-inflated however all residuals tests are normal other than that. I tried running it with zeros excluded and it only changed interpretation slightly with developed land (smaller effect size, though still relatively large, and non significant with zeros excluded)

I also considered using avg_productivity_all (which is an average including years where productivity was 0) but I don't think it's as good a representation and also was zero inflated in a glm with a poisson distribution

```{r productivity models}
hist(mod_phys$avg_productivity)
mean(mod_phys$avg_productivity, na.rm = TRUE) #1.555
var(mod_phys$avg_productivity, na.rm = TRUE) #1.550

hist(log10(mod_phys$avg_productivity))

mod.data.p<-mod_phys #%>% 
  #filter(avg_productivity>0)

#can't use poisson distributions because they're non-integer values
mp1<-lm(avg_productivity ~ BeachArea_km + beach_width + Length_km + parkArea + bufferPop_km2 + Border_road + tot_area_km_Beach + developedLand + Management + Region, data = mod.data.p)

mp2<-lm(avg_productivity ~ BeachArea_km + beach_width + Length_km + parkArea + bufferPop_km2 + Border_road + tot_area_km_Beach + developedLand + Region, data = mod.data.p)

summary(mp1)
summary(mp2)

AIC(mp1,mp2) #mp2 AIC = 196.6, mp1 AIC = 201.7

#exclusing zero values
mp3<-lm(avg_productivity ~ BeachArea_km + beach_width + Length_km + parkArea + bufferPop_km2 + Border_road + tot_area_km_Beach + developedLand + Region, data = (mod.data.p %>% filter(avg_productivity>0)))

summary(mp3)
```


```{r model validation prod}
simOutput <- simulateResiduals(fittedModel = mp2, plot = T)#okay

#tested zero inflation because we have a lot of 0 presence observations
testZeroInflation(simOutput) #problem
testDispersion(simOutput, type = "PearsonChisq", alternative = "greater") #n.s.

#excluding zero values
simOutputz <- simulateResiduals(fittedModel = mp3, plot = T)#okay

#tested zero inflation because we have a lot of 0 presence observations
testZeroInflation(simOutputz) #okay
testDispersion(simOutputz, type = "PearsonChisq", alternative = "greater") #n.s.
```


```{r spatial model effect size}
sjPlot::plot_model(mp2, show.values = TRUE, sort.est = TRUE, value.offset = .3)

#with no zero values effect size trends are the same however developed land is no longer significant
sjPlot::plot_model(mp3, show.values = TRUE, sort.est = TRUE, value.offset = .3)

```


### Human Models

#### Breeding Success
This model is A-OKAY
```{r logistic regression}
mod.data<-mod_human %>% 
  select(BeachID:Year, signs_small:vehicles, dogs_unleashedTotal:dogs_all, peopleIntensity, County, breedingSuccess) %>% 
  na.omit()

mb1<-glmer(breedingSuccess ~ signs_small + signs_large + dogs_all + peopleIntensity + dogs_unleashedTotal + vehicles + fencedNests + (1|County) + (1|BeachID), data = mod.data, family = binomial, control = glmerControl(optimizer = "bobyqa"))
#allFit(mb1) #bobyqa

summary(mb1) #AIC 216.2

# the significance is not different and this creates minor (n.s.) issues with the residuals in validation
mb2<-glmer(breedingSuccess ~ signs_small + signs_large + dogs_all + peopleIntensity + vehicles + dogs_unleashedTotal +  fencedNests + (1|County) + (1|BeachID), data = mod.data, family = binomial, control = glmerControl(optimizer = "bobyqa"))

summary(mb2) #AIC = 216.7
#AIC basically equivalent I will sue the vehicle model
```


```{r model validations}
simOutput <- simulateResiduals(fittedModel = mb1, plot = T) #some quantile abnormality but n.s.

testDispersion(mb1, type = "PearsonChisq", alternative = "greater") #okay, only looked for overdispersion
testZeroInflation(mb1) #okay
```


```{r plotting effect}
sjPlot::plot_model(mb2, show.values = TRUE, sort.est = TRUE, value.offset = .3)
```


#### Productivity model
```{r productivity human model}
mod.data.f<-mod_human %>% 
  select(BeachID:Year, signs_small:vehicles, dogs_unleashedTotal:dogs_all, peopleIntensity, County, Productivity, fledglingSuccess) %>% 
  na.omit()

mf1<-glmer(fledglingSuccess ~ signs_small + signs_large + dogs_all + peopleIntensity + dogs_unleashedTotal +  fencedNests + (1|County) + (1|BeachID), data = mod.data.f, family = binomial, control = glmerControl(optimizer = "Nelder_Mead"))
#allFit(mf1) #Nelder_Mead

summary(mf1) #AIC = 278.6

mf2<-glmer(fledglingSuccess ~ signs_small + signs_large + dogs_all + vehicles + peopleIntensity + dogs_unleashedTotal +  fencedNests + (1|County) + (1|BeachID), data = mod.data.f, family = binomial, control = glmerControl(optimizer = "Nelder_Mead"))
#allFit(mf1) #Nelder_Mead

summary(mf2) #AIC = 279.3

#Aic almost the same will use vehicle model
```

model okay
```{r fledgling validation}
simOutput <- simulateResiduals(fittedModel = mf2, plot = T) # good

testDispersion(mb1, type = "PearsonChisq", alternative = "greater") #okay
testZeroInflation(mb1) #okay
```


```{r effect size plots }
sjPlot::plot_model(mf2, show.values = TRUE, sort.est = TRUE, value.offset = .3)
```

## Figures

```{r tables}
beachTable<-data.frame(mod_phys$BeachID) %>% 
  rename(Beach = mod_phys.BeachID)
beachTable$SpatialModel<-"x"
beachTable$TemporalModel<-ifelse(beachTable$Beach %in% mod_human$BeachID, "x", "")
beachTable$Region<-mod_phys$Region
beachTable<-left_join(beachTable, mod_human[,c(1,23)], by = c("Beach"="BeachID")) %>% 
  unique() %>% 
  mutate(County = replace(County, Beach == "James", "Pictou")) %>% 
  mutate(County = replace(County, Beach == "Bowen Island", "Pictou")) %>% 
  mutate(County = replace(County, Beach == "Stoney (Lawrencetown Head)", "Halifax")) %>% 
  mutate(County = replace(County, Beach == "Cranberry Pond", "Queens")) %>% 
  mutate(County = replace(County, Beach == "Harbour Breeze, Port Joli", "Shelburne")) %>%
  mutate(County = replace(County, Beach == "Dominion (Lingan)", "Cape Breton")) %>% 
  mutate(County = replace(County, Beach == "Ogdens Pond", "Antigonish")) %>% 
  filter(!is.na(County)) %>% 
  arrange(Region, County) %>% 
  select(Region:County, Beach, SpatialModel, TemporalModel)

#shows which sites are in each model
write.table(beachTable)

#model tables, edit names in word
sjPlot::tab_model(mp2, m2, show.aic = TRUE)
sjPlot::tab_model(mb2, mf2, show.aic = TRUE)

#save and join together in word
sjPlot::tab_model(mf2, show.aic = TRUE)
sjPlot::tab_model(mb2, show.aic = TRUE)

#number of observations under 2 BP
sum(mod_phys$max_BP < 2)/sum(mod_phys$max_BP)
```


```{r phys mods eff size}
sjPlot::plot_model(m2, show.values = TRUE, value.offset = .3, ylim = c(0.1,1000))
sjPlot::plot_model(mp2, show.values = TRUE, value.offset = .3)

#saved 7x8 landscape as pdf
plot_models(mp2, m2, m.labels = c("Productivity", "Breeding Pairs"), show.p = TRUE, p.shape = FALSE, show.values = TRUE, spacing = .7, dot.size = 1.2, value.size = 3, axis.labels = c("Region","Developed Land","Foraging Area","Border Road","Part Border Road","No Border Road","Population Density","Parking Area","Beach Length","Beach Width","Beach Area"))

#this is worth discussing!
ggplot(mod_phys, aes(tot_area_km_Beach, developedLand, color = Region)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Foraging Area", y = "Developed Land") + scale_color_discrete(labels = c("Northern NS", "Southern NS"))
```


```{r human mods eff size}
sjPlot::plot_model(mf2, show.values = TRUE, value.offset = .3, ylim = c(0.1,1000))
sjPlot::plot_model(mb2, show.values = TRUE, value.offset = .3)

#do this for both, was going to do in ggplot but conversion of eff size for the other models would be a pain 
#(saved 7x8 landscape)
plot_models(mf2, mb2, m.labels = c("Productivity", "Breeding Pairs"), show.p = TRUE, p.shape = FALSE, show.values = TRUE, spacing = .7, dot.size = 1.2, value.size = 4.5, axis.labels = c("Fenced Nests","Mean Unleashed Dogs","Visitor Intensity","Vehicles","Mean Total Dogs","Large Signs","Small Signs"))
```


reformatting data to add location data to productivity data so I can potentially map some of this data in QGIS
```{r for Mapping, eval = FALSE}
centroids<-read.csv(file = '../Data/map.info.csv') %>% 
  select(Name,BeachArea, xcoord:ycoord) %>% 
  rename(BeachID = Name)

y<-centroids 
x<-mod_phys

#Make the location data have correct names
x<-stringdist_join(x, y, 
                by = "BeachID",
                mode = "left",
                ignore_case = FALSE, 
                method = "jw", 
                max_dist = 99, 
                distance_col = "dist") %>%
  group_by(BeachID.x) %>%
  slice_min(order_by = dist, n = 1) %>% 
  unique %>% 
  select(1,28,32)

colnames(x)<-c("BeachID", "correction", "dist")

#manually reassigning mismatches for fuzzy
x[x$BeachID == "Goose Haven", 2]<- "Port Joli (Goose Haven)"
x[x$BeachID == "White Point Beach", 2]<- "Gull Island Beach"

map_df<-full_join(centroids, x, by = c("BeachID" = "correction")) %>% 
  filter(!is.na(BeachID.y)) %>% 
  rename(BeachID_original = BeachID, BeachID = BeachID.y) %>% 
  select(-dist) %>% 
  left_join(mod_phys)
#write.csv( map_df, file = '../Data/map_phys_data.csv')
```

## End