---
title: "Beach Names"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

trying to subset the polygons to only work with NS beaches

Notes:
There's a note in the beachLength excel sheet saying the beach names conform with fed recovery data and spatial data so use these names (however I found a typo in the two Keji sites - no space between Keji and NP which was fixed manually in csv I think)

*No polygons/coastline present for:* 'Fox Bar', 'Harbour Breeze, Port Joli', 'Hirtles', 'Inverness Beach', 'Roaring Bull Point', 'White Point Beach', and 'Ogdens Pond'

*ALSO no polygons/coastline:* 'Chance Harbour', 'Goose and Burks Point' (Unless this is 'Goose (Indian) Point'), and 'Taylor Head Bay'

James & Bowen Island are grouped in BRC, but separate polygon sites.

*Coastline measurement only missing:* 'Goose Haven' and 'Gull Island Beach'

```{r}
beaches<-read.csv(file = '../Data/beaches polygon.csv') #polygon file with just columns, selected beaches within NS provincial boundary in QGIS
beaches$province<-NULL
beachLength<-read.csv(file = '../Data/beachLength.csv') #beach length csv from BRC

#creating a list of the correctly spelled beaches
beachNames<-beachLength$Critical.Habitat.Site.Name..Note..these.names.conform.with.fed..Recovery.docs.and.spatial.data.

#adding beaches in NS not in the beachLength doc but we have polygons for
beachNames<-append(beachNames, c("James", "Bowen Island", "Goose Haven", "Gull Island Beach")) 
beaches$province<-ifelse(beaches$Name %in% beachNames, "NS", "other")

summary(as.factor(beaches$province)) #56/61 beaches listed in beachLength, 5 polygon beaches missing and additionally 2 beaches missing polygons and coastline measurements

#double check we don't have polygons for these 5 beaches
missing<-setdiff(beachLength$Critical.Habitat.Site.Name..Note..these.names.conform.with.fed..Recovery.docs.and.spatial.data., beaches$Name) # missing coastline length and polygons

#this is all the beach polygons, not just the NS beaches to check I didn't accidentally exclude them
atlbeaches<-read.csv(file = '../Data/atl beaches polygons.csv')
atlbeaches %>% 
  filter(atlbeaches$Name %in% missing) # none so the polygons are really missing and need to be added

```
I can go into data sheets and find the GPS coordinates to digitize the 7 missing beaches


