---
title: "ProcessingData"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

To Do:
-get a number for each beach that shows the amount of land within the buffer (exclude sea) (DONE)
-finalize buffer size
-process landuse data (DONE)
-reformat with physical data (DONE)
-process survey data into format for analysis
-visualize
-pca of survey variables (not doing)


percentage of the landuse categories is based off land area present within the buffer, not including ocean or water bodies within.

```{r big dataset}
beachPhysW<-beachPhys %>% 
  pivot_wider(names_from = landclass, values_from = c(tot_area_km, tot_area, percentage)) %>% 
  select(-18,-23,-25,-28)

beachData<-full_join(prodData, beachPhysW[,c(1:6,8,10:13,15,22:24)], by = c('BeachID')) %>% 
  distinct() %>% #keeps only percentage for landuse and total_km for 2km beach buffer
  rename(Length_km = Length_km.y, Region = Region.y) %>% 
  select(-Length_km.x, -Region.x)


x<-piplSurvey %>% 
  summarise_all(funs(sum(is.na(.))))%>%
  pivot_longer(!BeachID, names_to = "variable", values_to = "count_na") %>% 
  select(2:3)

x %>% 
  filter(count_na < 1000)
```
Based on the above I think the main variables to include are:
dogs_leashedStart & dogs_unleashedStart (20,21)
signs_small & signs_large (25,26)
fencedNests & vehicles & litter (28,37,39)
engagedPeople & engagedPeople_dogs (33,34)
peopleTotal & dogs_leashedTotal & dogs_unleashedTotal (40,41,42)

columns: c(20:21,25:26,28,33:34,37,39:42)

### Exploration
I messed around with prcomp but I'm getting an error about missing/infinite values even when omitting NAs, which reminded me nmds with bray curtis is better for null values which we have
nmds isn't really suitable for large datasets though, so I'm moving on to other exploration/model prep

```{r landuse plots}
ggplot(beachPhys, aes(tot_area_km, color = landclass))+ geom_histogram()+facet_wrap(~landclass, scales = "free")

ggplot(beachPhys, aes(landclass, tot_area_km, colour = landclass))+ geom_boxplot()+facet_wrap(~Region)
```


```{r exploration}
piplSurvey[,c(18:21,25:26,28,33:34,37,39:42)]<-sapply(piplSurvey[,c(18:21,25:26,28,33:34,37,39:42)],as.numeric)
# count data so it isn't surprising we have left skew, probably poisson dist.
lapply(prodData[,c(4:8)], FUN = hist)
lapply(piplSurvey[,c(11:13,15:16)], FUN = hist) # pipl presence histos
lapply(piplSurvey[,c(20:21,25:26,28,33:34,37,39:42)], FUN = hist) #survey variables we want to include
lapply(beachPhysW[,c(3,7:21)], FUN = hist) #length and land buffer area are normal-ish, all others have left skew

#assessing which beaches have no pairs ever according to our data (prodData is CWS data)
# I could make this into a list for running analyses without these no pipl beaches
prodData %>% 
  group_by(BeachID) %>% 
  summarize(pairs = sum(YE_pairs), present = (sum(YE_pairs > 0)), total = n(), ratio = (sum(YE_pairs > 0)/n())) 
# 17 beaches with no pairs
# 11 beaches always have pairs, another 6 are 80% or above

```


```{r breeding boxplots}
boxplot(beachData[,c(5,8)]) #pairs has outliers, mostly crows neck
boxplot(beachData[,c(24:26)]) #some outliers in road and urban (probably near halifax?)
```

*Variables kept for possible inclusion in human models*
-dogs_leashedStart & dogs_unleashedStart (20,21)
-signs_small & signs_large (25,26)
-fencedNests & vehicles & litter (28,37,39)
-engagedPeople & engagedPeople_dogs (33,34)
-peopleTotal & dogs_leashedTotal & dogs_unleashedTotal (40,41,42)

Create a categorical variable for models to summarize multiple survey results
Using BRC class thresholds (notes in obsidian and in code)
```{r binning survey data}
#variables that have enough entries and are relevant
surveyVar<-c("dogs_leashedStart", "dogs_unleashedStart", "signs_small", "signs_large", "fencedNests", "vehicles","litter", "engagedPeople", "engagedPeople_dogs","peopleTotal","dogs_leashedTotal", "dogs_unleashedTotal","dogs_all")

piplSurvey$dogs_leashedTotal<-as.numeric(piplSurvey$dogs_leashedTotal)
piplSurvey$dogs_unleashedTotal<-as.numeric(piplSurvey$dogs_unleashedTotal)

piplSurvey$dogs_all<-ifelse((is.na(piplSurvey$dogs_unleashedTotal)&is.na(piplSurvey$dogs_leashedTotal)), NA, rowSums(piplSurvey[,c("dogs_leashedTotal", "dogs_unleashedTotal")], na.rm=TRUE))

#group by survey year
BRC_survey<- piplSurvey %>% 
  select(c(BeachID, year) | one_of(surveyVar)) %>% 
  mutate(across(one_of(surveyVar), as.numeric)) %>% 
  group_by(BeachID, year) %>% 
  summarize(across(c(one_of(surveyVar)), ~ mean(.x, na.rm = TRUE))) %>% 
  rename(Year = year)

#remove beaches with less than 4 surveys 
summary(as.factor(BRC_survey$BeachID))
lessSurvey<-c("Cranberry Pond","Green Bay", "Harbour Breeze, Port Joli", "Hawk Point & The Hawk", "Dominion (Lingan)", "Round Bay", "Roseway", "Stoney (Lawrencetown Head)", "Bowen Island", "Goose & Burks Point", "Goose and Burks Point")

BRC_survey<-BRC_survey %>% 
  filter(!BeachID %in% lessSurvey)

#with BRC classifications
BRC_survey$vehicles_ind<-ifelse(BRC_survey$vehicles>0.5, "poor", ifelse(
  BRC_survey$vehicles==0, "great", "fair"))
BRC_survey$offleash_ind<-ifelse(BRC_survey$dogs_unleashedTotal>1, "poor", ifelse(
  BRC_survey$dogs_unleashedTotal==0, "great", "fair"))

#add length column from beachPhys to calculate beachgoer intensity, create total dogs (sum off and on leash total) for percent off leash
BRC_survey$dogs_leashedPercent<-ifelse(
  is.na(BRC_survey$dogs_leashedTotal), NA, BRC_survey$dogs_leashedTotal/BRC_survey$dogs_all)
BRC_survey$dogRespect_ind<-ifelse(BRC_survey$dogs_leashedPercent<.5, "poor", ifelse(
  BRC_survey$dogs_leashedPercent>.80, "great", "fair")) #NOTE: I changed this from >1 (in BRC) to >.8 for great

BRC_survey<-left_join(BRC_survey, beachPhysW[,c(1,3)], by = 'BeachID')
BRC_survey$peopleIntensity<-BRC_survey$peopleTotal/BRC_survey$Length_km
BRC_survey$intensity_ind<-ifelse(BRC_survey$peopleIntensity>5, "poor", ifelse(
  BRC_survey$peopleIntensity<2, "great", "fair"))

BRC_survey[is.na(BRC_survey)]<-NA #changing all NaN to NA
```

### Model dfs

two models
one is year specific with human variables (survey data)
  -use the CWS pipl productivity as response variable
  -only include beaches surveyed sufficiently
one is beach specific and has all physical data (beach characteristics, landuse etc)
  -use total breeding pairs over surveyed period?
  -include more beaches? (as long as we have the CWS and phys data)

Maybe a model with only beaches that have pipl?
  - using relevant factors as informed by the other two models?

```{r prep for models}
mod_phys<-beachData %>% 
  filter(!is.na(YE_pairs)) #removes a few empty obs
mod_phys$County<-as.factor(mod_phys$County)
mod_phys$Region<-as.factor(mod_phys$Region)
mod_phys$Year<-as.factor(mod_phys$Year)
mod_phys$BeachID<-as.factor(mod_phys$BeachID)

#rescale BeachArea
mod_phys$BeachArea_km<-mod_phys$BeachArea/1000000

#format Management so it can be used more clearly
mod_phys$Management_raw<-mod_phys$Management #don't lose the OG
mod_phys$Management<-as.factor(ifelse(grepl(",", mod_phys$Management_raw), "Mixed", mod_phys$Management_raw))

#one developed landuse variable for simplicity
mod_phys$developedLand<-mod_phys$percentage_Agriculture + mod_phys$percentage_Urban + mod_phys$percentage_Road
```


```{r prep for human model}
prodData$Year<-as.character(prodData$Year)
mod_human<-left_join(BRC_survey, prodData[,1:9])
```


### End




